### 2004-05-17 Documentation/part2.tex
*** empty log message ***

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
Note that the energy environment will be described in the next section.

@@

The energy resolution function $R^c(E,E')$ has been already introduced in 
\Sec~\ref{sec:channel}, where a definition
has been given in \eq~(\ref{eq:e_res}). Instead of using
 \eq~(\ref{eq:e_res}) directly, we apply a slightly different
definition of the post-smearing efficiencies $\epsilon(E')$. 
In general, $\epsilon(E')$ has to be
determined by means of a Monte Carlo simulation of the experiment. 
This usually involves a binning of the simulated events in the 
reconstructed energy $E'$. Therefore, one simplify \eq~(\ref{eq:events_bin}) by
\begin{equation}
\label{eq:post_smearing}
\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \, R^c(E,E') \, \epsilon^c(E') = \hat\epsilon_i^c \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
Here the  $\hat\epsilon_i^c$ are the 
``post-smearing'' efficiencies, which will be set within the corresponding {\tt channel} environment (see below).
From \eq~(\ref{eq:events_bin}) it is obvious that the integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters. We define
the ``bin kernel'' $K_i^c$ for the $i$th bin as
\begin{equation}
\label{eq:kernel}
K_i^c(E) \equiv \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
With this definition, \eq~(\ref{eq:events_bin}) can be re-written as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int\limits_0^\infty dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}

@@

Before we come to the calculation algorithms, it is useful to understand
the general evaluation algorithm. As it is illustrated in \figu{energies}, 
\GLOBES\ uses several levels with respect to the energy ranges:
\begin{description}
\item[Sampling point level]
 This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points. The energy scale is the actual incident neutrino energy $E$. Note that the sampling range should exceed the energy bin range by about the energy resolution error, since these sampling points need to be evaluated by the algorithm. However, this is only
necessary to avoid aliasing effects if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
For a manual definition of the sampling points (depends on algorithm), use
\begin{quote}
{\tt
\$sampling\_points = 20\\
\$sampling\_min =          4.0\\
\$sampling\_max =         50.0
}
\end{quote}
for equidistant sampling points. Arbitrary sampling points can be specified analogously to the bins (see below).
\item[Bin level] This level is determined by the experiment. Note than energy
bin sizes much smaller than the energy resolution will not improve the results. The energy bin range and the number of energy bins do always have to be specified. For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.

In order to define a range between $E_\mathrm{min}$
and $E_\mathrm{max}$ divided by a certain number of equidistant bins,
use
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$bins = 20
}
\end{quote}
For arbitrary bins, use $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$:
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$binsize = \{  15.0 , 5.0 , 20.0, 6.0 \} 
}
\end{quote}
The number of bins will be automatically computed by \GLOBES . Note that the
bin sizes have to add up to the energy range {\tt \$emax}-{\tt \$emin}.
%
It is also
straightforward to compute the midpoint of each bin $E_i$. This information
has to be provided by the user ???? HOW ????. 
\item[Analysis level]
On the analysis level, an energy window can be defined within each rule. For
details, see next chapter.
\end{description}
In general, the energy smearing happens between the sampling point and bin levels, which means that the energy smearing matrix will have {\tt \$sampling\_points} columns and {\tt \$bins} rows.

As illustrated in the figure, an interesting feature in combination with the
channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the convolution in \eq~(\ref{eq:post_smearing}) is done. If they are introduced before, 
we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always afunction of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points (see \Sec~\ref{sec:energy}). In same cases, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment uses $10$ sampling points and $5$ bins.

In the following subsections we will define the energy resolution function.
All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}.
\begin{quote}
  {\tt energy(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}

%% WW: CONTINUE HERE

\subsection{Bin based automatic energy smearing}

First we discuss the builtin algorithms. The simplest, fastest and
least accurate is algorithm A. The key idea is to choose a set of assumptions
which allow to evaluate the integrand of equation~\ref{eq:simple_int}
at fixed values of the energy $E$ and to use the same number of points $N$ in
the calculation as there are bins. Thus equation~\ref{eq:simple_int} is reduced
to
\begin{equation}
\label{eq:algo_one}
n_i^c=N/L^2 \Delta E_i \sum_{j=1}^N \quad  \Phi^c(E_j)\,
P^c(E_j)\,
\sigma^c(E_j)\,
K_i^c(E_j)\,.
\end{equation}
The advantage of using equation~\ref{eq:algo_one} is that all factors
independent of the oscillation parameters have to be evaluated only at 
values of $E$ which are known in advance, \ie\ they can be put 
into a look-up table. Another advantage is that also the probability
has to be evaluated only at previously known values of the energy, which
makes it possible to compute the transition amplitudes for all channels
in one step. The key assumption is that all involved factors are piece-wise
constant, \ie\ they do not change within one bin. This assumption seems to be
very restrictive, note however whenever one is analyzing simulated 
data\footnote{simulated with the same algorithm} the errors
made by violating this assumption nearly cancel between the simulated 
data and the fitted data. This algorithm is the fastest which \GLOBES\ offers
and is suitable whenever there are no fast oscillations. Otherwise severe
aliasing effects may occur which usually lead to a failure of one of the 
minimizing routines. This algorithm is selected by
\begin{quote}
{\tt \tb @type = 1}
\end{quote}
The computation of the bin kernel $K_i^c$ is performed
by \GLOBES. To this end it requires  the number of bins
 $N_\mathrm{bins}$
and the minimum $E_\mathrm{min}$ and maximum energy $E_\mathrm{max}$. 
Furthermore following parameterization
for $R^c(E,E')$ is used
\begin{equation}
R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}}
\end{equation} 
There are several energy resolution functions available and by default
{\tt \#standard} is used. They can be chosen
by
\begin{quote}
{\tt \tb @sigma\_function = \#standard}
\end{quote}
where the energy resolution function named {\tt \#standard} is defined by
\begin{equation}
\label{eq:sigma_e}
\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\,.
\end{equation}
The parameters $\alpha, \beta$ and $\gamma$ are provided by the user.
\begin{quote}
{\tt \tb @sigma\_e = \{ 0.15, 0.0, 0.0\}}
\end{quote}
The other possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta}
which has only one parameter $\alpha$ and is defined by
\begin{equation}
\sigma(E)= \left\{\begin{array}{cl}
 \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
x>1.8\cdot10^{-3}\\
5\cdot10^{-5} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
\end{array} \right.
\end{equation}

In the actual implementation 
the sum in equation one only extends over those $E_j$ where $K(E_j)$ 
is above a certain threshold, by default $10^{-7}$. This threshold is 
defined at compile time. Thus a complete energy resolution definition of this 
type looks like
\begin{quote}
{\tt energy(\#name)<\\
\tb @type = 1\\
\tb @sigma\_function = \#standard\\
\tb @sigma\_e = \{ 0.15 ,0.0 ,0.0\}\\
>
}
\end{quote}

@@

@@

Finally the user has the possibility to directly specify the smearing data.
The format requires to provide $k_l^i$ and $k_u^i$ for each bin, the number of
bins, the number of sampling points, the lower and upper limit of the sampling
range. Furthermore the numbers $K_i(E_j)^c$ have to provide for all values
in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
equation~\ref{eq:algo_one}. Thus a complete example for this case would be
\begin{quote}
{\tt
\$simbins = 20\\
\$simtresh =          4.0\\
\$simbeam =         50.0\\
}

{\tt 
energy(\#name)<\\
\tb @energy =   \{0,2, 0.8634265, 0.0682827,     4e-06\}:\\
\tb\tb \{0,4, 0.1507103, 0.6965592, 0.1507103,   0.00101,     1e-07\}:\\
\tb\tb $\ldots$\\
>
}
\end{quote}

>>>>

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.

@@

The energy resolution function $R^c(E,E')$ has been already introduced in 
\Sec~\ref{sec:channel}, where a definition
has been given in \eq~(\ref{eq:e_res}). Instead of using
 \eq~(\ref{eq:e_res}) directly, we apply a slightly different
definition of the post-smearing efficiencies $\epsilon(E')$. 
In general, $\epsilon(E')$ has to be
determined by means of a Monte Carlo simulation of the experiment. 
This usually involves a binning of the simulated events in the 
reconstructed energy $E'$. Therefore, one simplify \eq~(\ref{eq:events_bin}) by
\begin{equation}
\label{eq:post_smearing}
\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \, R^c(E,E') \, \epsilon^c(E') = \hat\epsilon_i^c \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
Here the  $\hat\epsilon_i^c$ are the 
binned ``post-smearing'' efficiencies, which will be set within the corresponding {\tt channel} environment (see below).
From \eq~(\ref{eq:events_bin}) it is obvious that the integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters. We define
the ``bin kernel'' $K_i^c$ for the $i$th bin as
\begin{equation}
\label{eq:kernel}
K_i^c(E) \equiv \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
With this definition, \eq~(\ref{eq:events_bin}) can be re-written as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \,
\hat\epsilon_i^c \, \int\limits_0^\infty dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E) \, K_i^c(E)\,}_{f(E)}. 
\end{equation}

@@

Before we come to the calculation algorithms, it is useful to understand
the general evaluation algorithm. As it is illustrated in \figu{energies}, 
\GLOBES\ uses several levels with respect to the energy ranges:
\begin{description}
\item[Sampling point level]
 This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points. The energy scale is the actual incident neutrino energy $E$. Note that the sampling range should exceed the energy bin range by about the energy resolution error, since the corresponding sampling points have to be evaluated by the algorithm. 
However, this exceeding is only
necessary to avoid aliasing effects, if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
For a manual definition of the sampling points (depends on algorithm), use
\begin{quote}
{\tt
\$sampling\_points = 20\\
\$sampling\_min =          4.0\\
\$sampling\_max =         50.0
}
\end{quote}
for equidistant sampling points. Arbitrarily spaced sampling points can 
be specified analogously to the energy bins below. 
Note that the algorithm expects to 
have at least as many sampling points as energy bins, since otherwise 
the number of (unphysical) sampling constrains the experiment performance.
\item[Bin level] This level is determined by the experiment. Note than energy
bin sizes much smaller than the energy resolution will not improve the results. The energy bin range and the number of energy bins do always have to be specified. For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.

In order to define a range between $E_\mathrm{min}$
and $E_\mathrm{max}$ divided by a certain number of equidistant bins,
use
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$bins = 20
}
\end{quote}
For arbitrary bins, use $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$:
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$binsize = \{  15.0 , 5.0 , 20.0, 6.0 \} 
}
\end{quote}
The number of bins will be automatically computed by \GLOBES . Note that the
bin sizes have to add up to the energy range {\tt \$emax}-{\tt \$emin}.
\item[Analysis level]
On the analysis level, an energy window can be defined within each rule. For
details, see next chapter.
\end{description}
In general, the energy smearing happens between the sampling point and bin levels, which means that the energy smearing matrix will have {\tt \$sampling\_points} columns and {\tt \$bins} rows.

As illustrated in the figure, an interesting feature in combination with the
channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the integration in \eq~(\ref{eq:simple_int}) is done. If they are introduced before, 
we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always a function of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points. For some algorithms, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment uses $10$ sampling points and $5$ bins.

In the following subsections we will define the energy resolution function.
All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}.
\begin{quote}
  {\tt energy(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}
The individual parameters of the environment will be defined below and depend on the algorithm used.

\subsection{Bin-based automatic energy smearing}

This algorithm is the simplest of the built-in algorithms for the evaluation
of \eq~(\ref{eq:simple_int}). It is applicable to most of the
 accelerator-based experiments which can be simulated with \GLOBES .

The key idea is to use a ``flat'' model with sampling points identical to 
the bins (\cf, \figu{energies}). This is a good approximation as long as
\begin{itemize}
\item
 The number of bins is larger than the energy resolution.
\item
 The integrand in \eq~(\ref{eq:simple_int}) is small at the bin range limits
{\tt \$emin} and {\tt \$emax} (in order to avoid aliasing effects).
\item
 The neutrino oscillations are slow on a scale of the bin distance.
\end{itemize}
In this case, \eq~(\ref{eq:simple_int}) is reduced to
\begin{equation}
\label{eq:algo_one}
n_i^c=N/L^2 \, \sum_{j=1}^N \,  \Phi^c(E_j)\,
P^c(E_j)\,
\sigma^c(E_j)\,
K_i^c(E_j) \, \Delta E_j \,.
\end{equation}
The advantages of this algorithm are obvious: All factors
independent of the oscillation parameters have to be only evaluated once
at values of $E$ which are known in advance, which means that they can be put 
into a look-up table. In addition, the probability
has to be only evaluated at previously known values of the energy, which
makes it possible to compute the transition amplitudes for all channels
simultaneously. One assumption is that all involved factors are piece-wise
constant, \ie, they hardly change within each bin. This 
assumption seems to be very restrictive. Note, however, if one analyzes simulated data (which are simulated with the same algorithm), the errors
will cancel between the simulated and fitted data. 
%
Furthermore, we will demonstrate in the next subsection that the assumption
is rather good. ??? COMES THERE ???
%
This algorithm is the fastest which \GLOBES\ offers, and is selected by
\begin{quote}
{\tt \tb @type = 1}
\end{quote}
within the {\tt \#energy} environment.
The computation of the bin kernel $K_i^c$ is performed
by \GLOBES. Thus, it requires  the number of bins
 $N_\mathrm{bins}$
and the minimum $E_\mathrm{min}$ and maximum energy $E_\mathrm{max}$
for equidistant bins.

As far as the parameterization for the energy resolution function 
$R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
the Gaussian
\begin{equation}
R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
\end{equation} 
CHECK FACTOR OF TWO IN DENOMINATOR WITH SOFTWARE (IMPORTANT)!!! (NORMALIZATION IS OK!)
There are several energy resolution functions available, where by default
{\tt \#standard} is used:
\begin{quote}
{\tt \tb @sigma\_function = \#standard} 
\end{quote}
The energy resolution function {\tt \#standard} is defined by
\begin{equation}
\label{eq:sigma_e}
\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\, ,
\end{equation}
where the parameters $\alpha, \beta$ and $\gamma$ are provided by the user:
\begin{quote}
{\tt \tb @sigma\_e = \{0.15, 0.0, 0.0\}}
\end{quote}
Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
which only uses the parameter $\alpha$. It is defined by
\begin{equation}
\sigma(E)= \left\{\begin{array}{cl}
 \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
x>1.8\cdot10^{-3}\\
5\cdot10^{-5} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
\end{array} \right.
\end{equation}
In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only computed for the $E_j$'s where $K(E_j)$ 
is above a certain threshold, which is by default $10^{-7}$. 
This threshold is defined at the compiling time.

Eventually,  a complete energy resolution definition of this 
algorithm is, for example,
\begin{quote}
{\tt energy(\#name)<\\
\tb @type = 1\\
\tb @sigma\_function = \#standard\\
\tb @sigma\_e = \{0.15 ,0.0 ,0.0\}\\
>
}
\end{quote}

@@

THIS SUBSECTION WILL BE RE-WRITTEN BY PATRICK (WW)

@@

In some cases, one may use the output of a detector Monte Carlo simulation
directly. Then one can use manual energy smearing instead of one of the
automatic energy smearing algorithms.

The energy smearing matrix $K_{ij}$ has {\tt \$bins} rows and {\tt \$sampling\_points} columns, which are numbered from $0$ to {\tt \$bins}$-1$ or {\tt \$sampling\_points}$-1$. It is equivalent to the 
bin- and sampling-point-based kernel in \eq~(\ref{eq:kernel}):
\begin{equation}
K_{ij} = K_i^c(E) |_{E=E_j},
\label{equ:ematrix}
\end{equation}
where $E_j$ is the energy of the $j$th sampling point. In general, many of the entries in this matrix are zero, which means that it is convenient to evaluate the integrand in \eq~(\ref{eq:simple_int}) only at positions where
$K_{ij}$ is non-zero. The corresponding ``sampling range'' range of non-zero matrix entries  in $K_{ij}$ for the $i$th energy bin is defined to run from
column $k_l^i$ (``lower index'') to column $k_u^i$ (``upper index'').
An example for a smearing matrix is
\begin{equation}
K_{ij} =   \underbrace{ \left( \begin{array}{cccccccccc} 
a_{00} & a_{01} & a_{02} & a_{03} &  &  &  &  &  & \\
a_{10} & a_{11} & a_{12} & a_{13} & a_{14} &  &  &  &  &  \\
 & a_{21} & a_{22} & a_{23} & a_{24} & a_{25} &  &  &  &  \\
 &  & a_{32} & a_{33} & a_{34} & a_{35} & a_{36} &  &  &  \\
 &  &  & a_{43} & a_{44} & a_{45} & a_{46} & a_{47} &  &  \\
& & & \uparrow  & & \ddots & & \uparrow & & \\
& & & k_l^i & & & & k_u^i & & \\ 
\end{array}\right) }_{\mathtt{\$sampling\_points}  \, \, \mathrm{columns}} \quad \leftarrow \quad \footnotesize{\mathtt{\$bins} \, \, \mathrm{rows}} \, ,
\end{equation}
where the unshown entries are zero. Thus, the values of $K_{ij}$ have to be specified between $k_l^i$ and $k_u^i$ in the form $\{ k_l^i,k_u^i, K_{i \, k_l^i}, K_{i k_l^i+1} , \hdots , K_{i k_u^i} \}$:
\begin{quote}
{\tt 
energy(\#name)<\\
\tb @energy =   \{0,2, 0.8634265, 0.0682827,     4e-06\}:\\
\tb\tb \{0,4, 0.1507103, 0.6965592, 0.1507103,   0.00101,     1e-07\}:\\
\tb\tb $\ldots$\\
>
}
\end{quote}
Note that the sum of all entries in each {\em column} should be equal 
to unity,
since all of the incoming neutrinos should be assigned to energy bins. In many practical cases, however, the definition of the energy smearing can
lead to sums smaller than unity, such as in the case of truncated Gaussian
distributions. The sum of entries in each {\em row} is not defined, since the events might be unevently distributed into the energy bins
according to the energy resolution function.
